---
title: "get_troopdata"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{troopdata-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

library(ggplot2)
```

This page provides an overview for the `get_troopdata()` function, highlighting some of its potential uses.

First things first—let's load the `{troopdata}` package


```{r setup}
library(troopdata)
```

The troopdata package provides multiple functions to generate customizable datasets containing information on US military deployments and accompanying data. The `get_troopdata()` function represents the core of this package, providing customized data on US overseas troop deployments, specifically.


## Country-year data

The first function of this package is the `get_troopdata()` function. At its most basic this function returns a data frame of country-year troop deployment values for the selected time period, using the `startdate` and `enddate` parameters.

```{r example, warning=FALSE, echo = FALSE, message=FALSE}
library(troopdata)
## basic example code

# This example uses the basic function to gather total troop figure
example <- get_troopdata(startyear = 1990, endyear = 2020)

```


For users who want more refined data, the `host` argument and the `branch` arguments allow users to specify the set of host countries for which they would like data returned. This must be a single numerical value equal to a Correlates of War (COW) Project country code, a single character value equal to an ISO3C country code, or a vector of similar values. Note they must be consistent (i.e. they must all be numeric COW codes or ISO3C character codes).

For example, you can use a numeric vector of COW country codes:

```{r}

# Let's make the host selection more specific
hostlist <- c(200, 220)

example <- get_troopdata(host = hostlist, startyear = 1990, endyear = 2020)

head(example)

```

Or you can use a character vector of ISO3C codes.

```{r}

hostlist.char <- c("CAN", "GBR")

example.char <- get_troopdata(host = hostlist.char, startyear = 1970, endyear = 2020)

head(example.char)

```




```{r}

hostlist <- c(20, 200, 220)

example <- get_troopdata(host = hostlist, branch = TRUE, startyear = 2006, endyear = 2020)

head(example)

```


## Regional Summaries

Sometimes users are also interested in obtaining data on regional patterns in U.S. deployments. Now users can enter `host = "region"` to obtain regional sums of U.S. deployments. The `year` and `branch` arguments work as in the previous examples, allowing users to generate summed values for aggregate troop deployments or branch-specific deployment sums by region. These values can be plotted as in previous examples, too.

```{r}

regional.sum <- get_troopdata(host = "region", startyear = 1990, endyear = 2020) 

ggplot(regional.sum, aes(x = year, y = troops, color = region)) + 
  geom_line() +
  viridis::scale_color_viridis(discrete = TRUE) +
  labs(x = "Year",
       y = "Count",
       title = "Troop deployments by region",
       color = "Region")
  

```
