---
title: "get_builddata"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{get_builddata}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

```

This page provides an overview for the `get_builddata()` function, highlighting some of its potential uses.

First things first—let's load the `{troopdata}` package


```{r setup}
library(troopdata)
library(ggplot2)
```



## Basic Use

The `get_builddata()` function returns a data frame containing geocoded location-project-year military construction data. The basic arguments function the same as compared to the previous functions. The primary difference is that the data cover fiscal years 2000 through 2026, as published in the Department of Defense Comptroller Annual Report C-1 exhibits. Note there are also many observations included that contain amounts, but do not disclose location names or other information.

```{r build example}

hostlist <- c(200, 260, 211) # United Kingdom, Germany, Belgium (Gleditsch and Ward codes)

buildexample <- get_builddata(host = hostlist, startyear = 2008, endyear = 2019)

head(buildexample)


```


As with the base data you can build cool maps using the construction data. You can also size the points according to the amount of spending associated with a particular location, adding some additional details to maps and other figures.

```{r buildmap, echo = TRUE, warning = FALSE, message=FALSE, dpi = 400}

library(ggplot2)

map <- ggplot2::map_data("world")
basepoints <- troopdata::get_builddata(host = NA, startyear = 2009, endyear = 2019)


buildmap <- ggplot() +
  geom_polygon(data = map, aes(x = long, y = lat, group = group), fill = "gray80", color = "white", size = 0.1) +
  geom_point(data = basepoints, aes(x = longitude, y = latitude, size = toa_amount), color = "purple", alpha = 0.4) +
  coord_equal(ratio = 1.3) +
  theme_void() +
  theme(plot.title = element_text(face = "bold", size = 15)) +
  labs(title = "Locations of U.S. military construction spending, 2009-2019",
       size = "Spending Amount")


buildmap

```
