---
title: "get_troopdata"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{get_troopdata}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

```

This page provides an overview for the `get_troopdata()` function, highlighting some of its potential uses.

First things first—let's load the `{troopdata}` package


```{r setup}
library(troopdata)
library(ggplot2)
library(tidyverse)
library(viridis)
```

The `troopdata` package provides multiple functions to generate customizable datasets containing information on US military deployments and accompanying data. The `get_troopdata()` function represents the core of this package, providing customized data on US overseas troop deployments, specifically.


## Country-year data

The first function of this package is the `get_troopdata()` function. At its most basic this function returns a data frame of country-year troop deployment values for the selected time period, using the `startyear` and `endyear` parameters. Here are the values for the United States from 1990 to 2020.

```{r example, warning=FALSE, echo = FALSE, message=FALSE}
## basic example code
library(troopdata)
# This example uses the basic function to gather total troop figures for the United States from 1990 to 2020
example <- get_troopdata(host = "USA", startyear = 1990, endyear = 2020)

head(example)

```


For users who want more refined data, the there are a number of arguments that allow the user to further tailor the output to their needs.

The `host` argument allows users to specify the set of host countries for which they would like data returned. If you leave it out you get every location in the data, the United States included. This can be a vector of numerical values equal to a Gleditsch and Ward country code, a vector of character values equal to an ISO3C country code, or a vector of character values corresponding to full country names. Note that when supplying a vector of values they must be consistent and correspond to a single type of identifier at a time (i.e. they must all be numeric country codes, ISO3C character codes, country names, or region names). If one of the values does not match anything, the function returns the rest and gives a warning that names the value it could not match.

A note on the numeric codes: the data use the Gleditsch and Ward list, not the Correlates of War list. The two agree for most countries, but not all of them. Germany, for example, is 260 here and not 255. Territories and other locations that the Gleditsch and Ward list does not cover have codes of our own, such as 6 for Puerto Rico, 1002 for Greenland and 1008 for Guam. A numeric code that is not in the data returns an error rather than an empty data frame.

For example, you can use a numeric vector of country codes like this:

```{r}

# Let's make the host selection more specific
hostlist <- c(200, 220)

example <- get_troopdata(host = hostlist, startyear = 1990, endyear = 2020)

head(example)

```

Or you can use a character vector of ISO3C codes.

```{r}

hostlist.char <- c("CAN", "GBR")

example.char <- get_troopdata(host = hostlist.char, startyear = 1970, endyear = 2020)

head(example.char)

```

A few locations have no ISO3C code of their own and carry their parent's, so one code can return more than one location. `"GBR"` returns the United Kingdom and Akrotiri, `"PRT"` returns Portugal and the Azores, and `"UMI"` returns Wake, Midway and Johnston Islands. Each location comes back on its own row with its own country code, and a warning tells you when this happens. Add the rows up if you want a single figure for the ISO3C code.

Similarly, we can search for full country names:

```{r}
hostlist.names <- c("Canada", "United Kingdom")

example.names <- get_troopdata(host = hostlist.names, startyear = 1970, endyear = 2020)

head(example.names)

```

When searching for country names, the function will do its best to identify the correct country based on the character string that's included. This can include cases where fragments of country names are included and the function will try to return the correct country. 

```{r}

example.frag <- get_troopdata(host = "South Ko", startyear = 1970, endyear = 2020)

head(example.frag)

```


Finally, we can also search by region. Instead of inserting a country name or code into the host argument you can simply include character strings that represent regions. In these cases the function returns every country in the matching region or regions for the specified time period, one row per country and year, with the region in its own column. Nothing is added up across the countries of a region.

The regions are East Asia & Pacific, Europe & Central Asia, Latin America & Caribbean, Middle East & North Africa, North America, South Asia and Sub-Saharan Africa. (Antarctica and the "Afloat" location are in regions of their own.) Region names are matched as fragments, which calls for a little care:

- A fragment that matches one region works fine. `"Europe"` returns Europe & Central Asia.
- A fragment that matches several regions returns all of them, with a warning. `"Asia"` returns the countries of East Asia & Pacific, South Asia, and Europe & Central Asia.
- A fragment that is also part of a country name is read as a country, again with a warning. `"Africa"` returns South Africa and the Central African Republic, so ask for `"Sub-Saharan Africa"` instead.
- Regions and country names cannot be mixed in one call.

```{r}

region.list <- c("Europe & Central Asia", "East Asia & Pacific")

example.region <- get_troopdata(host = region.list, startyear = 1970, endyear = 2020)

head(example.region)

```



## Disaggregated Data

By default the `get_troopdata()` function returns the aggregate sum of active duty military personnel. But the original DMDC reports often include disaggregated figures, with separate counts for each branch of the military. The `branch` argument allows users to specify whether they would like to receive the aggregate sum of all branches or the disaggregated figures for each branch. This argument can take on two values: `TRUE` or `FALSE`, with the default being `FALSE`. 

```{r}

# Let's get the disaggregated data for the US deployments to Canada and the UK
hostlist <- c("Canada", "United Kingdom")

example.branch <- get_troopdata(host = hostlist, branch = TRUE, startyear = 1970, endyear = 2020)

head(example.branch)


```

In each case the `_ad` suffix on the variable name indicates "Active Duty" numbers for the given branch.

Note that the total does not necessarily equal the sum of the individual branches. The function returns the maximum annual value for each branch. In cases where there are quarterly values reported, the sum total may come from one quarter and the individual branch values may come from another quarter. The section on time periods below shows an example.

We can also include disaggregated data national guard and reserve personnel, as well as DoD civilians. The reports give these numbers from 2008 forward, so they show up as `NA` for earlier years. Setting `guard_reserve = TRUE` also returns `troops_all`, which is the active duty total plus the guard and reserve columns. Before 2008 it is the same as `troops_ad`.

```{r}

hostlist <- c("Canada", "United Kingdom")

example.branch <- get_troopdata(host = hostlist, branch = TRUE, startyear = 1970, endyear = 2020, guard_reserve = TRUE, civilians = TRUE)

head(example.branch)

```


## Time Periods

The most recent update also allows users to specify more fine grained temporal coverage. DMDC reports have historically been released on an annual basis, but in more recent years they have been released twice annually or even quarterly, and the `get_troopdata()` function allows users to specify whether they would like to receive the quarterly data or the annual data. The `quarters` argument allows users to specify whether they would like to receive the quarterly data or the annual data. This argument can take on two values: `TRUE` or `FALSE` with the default being `FALSE`. 

The data have one report per year through 2012, two for 2013 (September and December), and four per year from 2014 forward (March, June, September and December). The most recent year only has the reports that have been published so far.

If the user opts to return quarterly data, the function will return the month and quarter columns in addition to the year, and each row holds the figures from that one report. A country that a report does not list gets a 0 for that period rather than `NA`, so a zero can mean either that the report shows zero or that the country is not in it. Territories such as Guam or Bermuda are different: they only have a row for the periods in which a report lists them.

Here we use full country names. See! Neat!


```{r}

# Let's get the quarterly data for the US deployments to Canada and the UK
hostlist <- c("Canada", "United Kingdom")

example.quarters <- get_troopdata(host = hostlist, branch = TRUE, startyear = 2015, endyear = 2022, quarters = TRUE)

head(example.quarters)

```

### How the annual values are chosen

When a year has more than one report, the annual data return the **highest value reported in that year**. This is not an average, and it is not the value from one particular month. It is also done separately for each column, so the total can come from one report and a branch figure from another.

Japan in 2019 is a good example. Here are the four quarterly reports:

```{r japan-quarters, warning = FALSE}

japan.quarters <- get_troopdata(host = "Japan", branch = TRUE, startyear = 2019, endyear = 2019, quarters = TRUE)

# Put the rows in calendar order and keep the columns we want to compare
japan.quarters <- japan.quarters[order(japan.quarters$quarter), ]

japan.quarters[, c("countryname", "year", "month", "troops_ad", "army_ad", "navy_ad", "air_force_ad", "marine_corps_ad")]

```

And here is the annual row for the same year:

```{r japan-annual, warning = FALSE}

japan.annual <- get_troopdata(host = "Japan", branch = TRUE, startyear = 2019, endyear = 2019)

japan.annual[, c("countryname", "year", "troops_ad", "army_ad", "navy_ad", "air_force_ad", "marine_corps_ad")]

```

The annual `troops_ad` value is the December figure, the largest of the four. But the annual `army_ad` value is the June figure and the annual `navy_ad` value is the March figure, because those were the largest for each of those branches. This is why the branch columns in the annual data do not have to add up to `troops_ad`. If you need figures that all come from the same report, use `quarters = TRUE`.

Before 2013 there is only one report per year, so the annual value is the value from that report, or the outside figure described below where the report has none.

### Gaps in the reports

A few values do not come straight from a report. The big ones are:

- **1951 and 1952.** There are no reports for these years. For countries the totals move in equal steps between the 1950 and 1953 figures, and there are no branch figures. Most territories have no rows for these two years.
- **December 2022, March 2023 and June 2023.** The Army did not report any personnel figures for these quarters while it switched to a new personnel system. In the country data the Army figures for these quarters move in equal steps between the September 2022 and September 2023 reports, and the totals are those figures plus what the other services reported. The data for US states and the reports themselves are left as published, so the Army figures and the totals are `NA` there.
- **June 2023.** The June 2023 report has its civilian columns and some of its guard and reserve columns printed one row too low for part of the overseas list, so that Qatar's row shows Puerto Rico's numbers. The country data put these figures back on the right rows. The reports themselves (`reports = TRUE`) are left as published.
- **Afghanistan, Iraq, Kuwait and Syria.** The reports do not give the number of personnel deployed to these countries for much of the 2000s, and from December 2017 they leave out deployed personnel in Afghanistan, Iraq and Syria again. Where we could, we use figures from Kane's data or from public reporting instead (Afghanistan 2001-2005 and 2018-2020, Iraq 2003-2007 and 2018-2021, Kuwait 2003-2007, and Syria 2018-2021). These are totals only, so the branch columns are zero. After those years the values are whatever the reports show, which is only a handful of people and well below the actual US presence. For Iraq in 2018-2021 and Syria in 2021 the outside figure only appears in the June row of the quarterly data, so use the annual data for those.

The [Rebuild Notes](https://meflynn.github.io/troopdata/articles/01-rebuild-notes-vignette.html) article and the help page for `get_troopdata()` go through all of this in more detail.

## Personnel Afloat

The reports for 1950 through 2007 count personnel ashore and personnel afloat separately, and for some years they assign the afloat personnel to a country, usually the country of the nearest port. By default the country figures only include personnel ashore. The `afloat` argument lets you change that:

- `afloat = "exclude"` is the default. Country figures are personnel ashore only.
- `afloat = "include"` adds the afloat personnel assigned to a country to that country's totals.
- `afloat = "separate"` keeps the ashore figures as they are and returns the afloat personnel next to them in the `troops_afloat`, `navy_afloat` and `marine_corps_afloat` columns.

```{r greece-afloat, warning = FALSE}

greece <- get_troopdata(host = "Greece", startyear = 1958, endyear = 1962, afloat = "separate")

greece[, c("countryname", "year", "troops_ad", "troops_afloat", "navy_ad", "navy_afloat")]

```

There were about 2,000 personnel ashore in Greece in 1960, and more than 13,000 afloat who were counted there because that is where the fleet was on the day of the count.

Everyone afloat is also carried in a separate location named "Afloat" (country code 10200). It is not a country, so remember to leave it out if you are adding up countries. From September 2008 the reports count a ship's crew at its home port, as part of the country or state figure, so the argument makes no difference for 2008 and later.

## Data for US States

The package now incorporates data on the number of US service personnel in each US state. The functionality of the command works just as it does when users are looking to pull data for countries, with a couple of small changes. 

First, these data come from the DMDC reports beginning in 2008, so we don't have as much time to work with as we do with the country-level data. They cover the fifty states and the District of Columbia. Puerto Rico, Guam and the other territories are in the country-level data.

Second, users need to make sure they set `state_data = TRUE` in the function call. The data for US states work with different identifiers, so this argument needs to be set for the function to differentiate between queries looking at country data and queries looking at data on US states.

Finally, as mentioned above, the data for US states use different identifiers for the relevant geographic units. Users can search for data by the state's name using character strings, just like in the country-level data. Capitalization does not matter. A full state name returns just that state, so `"kansas"` gives you Kansas and `"Virginia"` does not also give you West Virginia, while part of a name such as `"Carolina"` returns every state that contains it. Users can also look to match US states by their numeric FIPS codes. These are useful if we're interested in generating things like maps using the state-level data.

The `branch`, `guard_reserve`, `civilians` and `quarters` arguments work the same way they do for countries, and annual values are again the highest value reported in the year. Two things to keep in mind. The Army did not report for December 2022, March 2023 and June 2023, and the state data leave the Army figures and the totals as `NA` for those quarters rather than filling them in, so the annual Army figures and totals for 2022 and 2023 come from the other quarters. And from December 2015 to December 2017 the reports list a large number of Navy personnel in no state at all, so the Navy figures for home-port states like Virginia and California are lower in those quarters.

```{r state-maps}
library(usmap)
library(ggplot2)

states <- get_troopdata(startyear = 2025, endyear = 2025, state_data = TRUE)

map.data <- usmap::us_map(regions = "states") %>% 
  dplyr::mutate(fips = as.numeric(fips)) %>% 
  left_join(states, by = c("fips" = "fipscode"))

ggplot2::ggplot(data = map.data) +
  geom_sf(aes(geometry = geom, fill = troops_ad)) +
  theme_void() +
  viridis::scale_fill_viridis(option = "turbo") +
  labs(fill = "Military Personnel")

```




## Reports

Finally, users may want to view the original DMDC reports that the data is drawn from. The `reports` argument allows users to specify whether they would like to receive the original DMDC reports that the data is drawn from. This argument can take on two values: `TRUE` or `FALSE` with the default being `FALSE`.

Users can specify the `host`, `startyear`, and `endyear` arguments as they would for the main function. The function will return a single data frame containing all of the original columns found in the DMDC reports upon which the data are based. The formatting and column names will be roughly consistent with the original reports, but the data will be filtered to only include the specified host countries and time period. 

The reports are returned row by row, as they were published, with nothing added up. This means a country code can show up on more than one row in the same period when a report lists parts of a country separately. Japan and the Ryukyu Islands (Okinawa), for example, are separate lines in the reports through 1973 and both come back when you ask for Japan. The country-level data add those rows together.

The `source` column in the data frame provides the month and the year that the data was drawn from. This allows the user to more easily track down the original DMDC report that the data was drawn from. The current and archived reports can be found here: [DMDC Reports](https://dwp.dmdc.osd.mil/dwp/app/dod-data-reports/workforce-reports)

Also note that if `reports` is set to `TRUE` then the user must also set the `quarters` argument to `TRUE`. 


